---
title: "Geocomputation with R chapter 2"
output: html_notebook
---


```{r}
devtools::install_github("Nowosad/spDataLarge")
```

```{r}
library(sf)          # classes and functions for vector data
library(raster)      # classes and functions for raster data
library(spData)        # load geographic data
library(spDataLarge)   # load larger geographic data
```

#2.2 vector data

##2.2.1 sf package

```{r}
vignette("sf1")
```

```{r}
nc <- st_read(system.file("shape/nc.shp", package="sf"))
```

```{r}
class(nc)
```

```{r}
print(nc[9:15], n = 3)
```

```{r}
par(mar = c(0,0,1,0))
plot(nc[1], reset = FALSE) # reset = FALSE: we want to add to a plot with a legend
plot(nc[1,1], col = 'grey', add = TRUE)
```


```{r}
nc[1,1]
```

forget about the rest of the vignette, I've seen it already....

```{r}
names(world)
```
```{r}
plot(world)
```

And you can treat a spatial object like a dataframe. So you can query it for things.

```{r}
summary(world["lifeExp"])
```

let's go off on a tangent and do something from ch. 2 Into Econometrics R:

```{r}
par(cex.axis=0.8)
boxplot(lifeExp ~ continent, data = world)
```


How to subset an sf:

```{r}
worldaf <- subset(world, continent == "Africa")
plot(worldaf)
```

Ha! I even figured out how to do that myself!

And let's do a dotplot of that:

```{r}
aa <- lm(formula = lifeExp ~ gdpPercap, data = worldaf)
summary(aa)
```

```{r}
plot(lifeExp ~ gdpPercap, data = worldaf)
```

Looks off. Let's log both:

```{r}
plot(log(lifeExp) ~ log(gdpPercap), data = worldaf)
```

Now regress this:

```{r}
aa <- lm(formula = log(lifeExp) ~ log(gdpPercap), data = worldaf)
summary(aa)
```


Are we sure that both are distributed normally?

```{r}
library(ggplot2)
```


```{r}
a <- ggplot(worldaf, aes(x = log(gdpPercap)))
a <- a + geom_density() 
a
```

```{r}
a <- ggplot(worldaf, aes(x = log(lifeExp)))
a <- a + geom_density() 
a
```

Neither is really normal. But whatever.


You can convert an sf object to an sp object easily:

```{r}
library(sp)
world_sp = as(world, Class = "Spatial")
```

##2.2.3 basic map making

You can join things together to make bigger things:

```{r}
world_asia = world[world$continent == "Asia", ]
asia = st_union(world_asia)
plot(asia)
```


Then plot asia on top of world:


```{r}
plot(world["pop"], reset = FALSE)
plot(asia, add = TRUE, col = "red")
```








