---
title: "Intro to Econometrics with R Ch. 1"
output: html_notebook
---

#1.3 - Basic Calculations

```{r}
pi
```

```{r}
exp(1)
```

Ooh, you can do logs in any base?

```{r}
log(10)
log10(10)
log2(10)
log(10, base = 77.1)
```

```{r}
77.1^0.5299264
```

#1.7 - install the Sciences Po Econometrics Package

```{r}
library(devtools)
```


```{r}
install_github(repo = "ScPoEcon/ScPoEconometrics")
```
```{r}
library(ScPoEconometrics)
packageVersion("ScPoEconometrics")
```


# 1.9 Creating a sequence

```{r}
seq(from = 1.5, to = 4.2, by = 0.1)
```

from, to, and by are optional:

```{r}
seq(1.5, 4.2, 0.1)
```

or use rep for repeat:

```{r}
rep("A", times = 10)
```

subsetting:

```{r}
x = c(1, 3, 5, 7, 8, 9)
x
```

leave out element 2:

```{r}
x[-2]
```

part of a vector:

```{r}
x[1:3]
x[c(1,3,4)]
```

subsetting with a vector of logical values:

```{r}
z = c(TRUE, TRUE, FALSE, TRUE, TRUE, FALSE)
x[z]
```

math operation on a vector

```{r}
x = 1:10  # a vector
x
x + 1     # add a scalar
2 * x     # multiply all elements by 2
```

then reverse a vector:

```{r}
rev(x)
```

make a matrix out of vectors using rowbind:

```{r}
x = 1:9
rev(x)
rep(1, 9)
rbind(x, rev(x), rep(1, 9))
```

column bind instead of row bind:

```{r}
cbind(col_1 = x, col_2 = rev(x), col_3 = rep(1, 9))
```

here btw is a transpose:

```{r}
t(cbind(col_1 = x, col_2 = rev(x), col_3 = rep(1, 9)))
```

This gives the inverse of a square matrix if it's invertible:

```{r}
x = c(3,0,3,0,2,2,1,5,5)
y = 9:1
(X = matrix(x, 3, 3))
(Y = matrix(y, 3, 3))
(Z <- solve(X))
X %*% Z
```

## 1.9.4 Arrays of higher dimensionality

You can create an array of arbitrary dimensionality N.

## 1.9.5 Lists

Lists are one-dimensional heterogeneous data structures (while vectors are homogenous).

# 1.10 Data frames

```{r}
path = system.file(package="ScPoEconometrics","datasets","example-data.csv")
example_data_from_disk = read.csv(path)
example_data_from_disk
```

```{r}
mtcars
```


use head to print the first n observations:

```{r}
head(mtcars,n=2)
```


str() will instead display the structure of the dataframe:

```{r}
str(mtcars)
```


this dataset has a help:

```{r}
?mtcars
```

You can get a vector of the variable names for a dataframe:

```{r}
names(mtcars)
```

Here's how to get the dimensions of a dataframe:

```{r}
dim(mtcars)
nrow(mtcars)
ncol(mtcars)
```

You can subset dataframes. Here's cars with >25mpg, and only 3 columns:

```{r}
mtcars[mtcars$mpg > 20, c("cyl", "disp", "wt")]
```

Or do it this way:

```{r}
subset(mtcars, subset = mpg > 25, select = c("cyl", "disp", "wt"))
```
 

#1.11 Looping

Here's a for loop:

```{r}
for (i in 1:5){
  print(i)
}
```

You can iterate thru anything, not just numbers:

```{r}
for (i in c("mangos","bananas","apples")){
  print(paste("I love",i))  # the paste function pastes together strings
}
```

```{r}
for (i in 2:3){
  # first nest: for each i
  for (j in c("mangos","bananas","apples")){
    # second nest: for each j
    print(paste("Can I get",i,j,"please?"))
  }
}
```

##1.11.4 functions

Here is a function:

```{r}
hello <- function(your_name = "Lord Vader"){
  paste("You R most welcome,",your_name)
  # we could also write:
  # return(paste("You R most welcome,",your_name))
}

# we call the function by typing it's name with round brackets
hello()
```

Just typing the function returns the definition to us.

```{r}
hello
```

That's not a vector, though:

```{r}
dim(hello)
```









