---
title: "Intro to Econometrics with R Ch. 2"
output: html_notebook
---

#2.1 Summary statistics

```{r}
library(ggplot2)
```

Let's use that mpg dataset

```{r}
mpg
```


```{r}
mpg$cty
```

We can get mean and median:

```{r}
mean(mpg$cty)
median(mpg$cty)
```

We can get variance, SD, interquartile range, minimum, maximum, and range:

```{r}
var(mpg$cty)
sd(mpg$cty)
IQR(mpg$cty)
min(mpg$cty)
max(mpg$cty)
range(mpg$cty)
```


#2.2 Plotting

You can plot very simple things using R's built-in plot facilities:

```{r}
hist(mpg$cty)
```


And we can pretty up a plot:

```{r}
hist(mpg$cty,
     xlab   = "Miles Per Gallon (City)",
     main   = "Histogram of MPG (City)", # main title
     breaks = 12,   # how many breaks?
     col    = "red",
     border = "blue")
```


For a categorical variable, you can use barplot:

```{r}
barplot(table(mpg$drv),
        xlab   = "Drivetrain (f = FWD, r = RWD, 4 = 4WD)",
        ylab   = "Frequency",
        main   = "Drivetrains",
        col    = "dodgerblue",
        border = "darkorange")
```


You can find out how many unique values are in each column:

```{r}
unique(mpg$drv)
```


You can use a boxplot to visualize a single numerical variable.

The box is the interquartile range, the whiskers are 1.5x IQR, the midline is the median, and the circles are outliers from the IQR.

```{r}
boxplot(mpg$hwy)
```


And you can use this to compare variables:

```{r}
boxplot(hwy ~ class, data = mpg)
```

Note it didn't print 2 classes:

```{r}
unique(mpg$class)
```


The text was too big. 

We can fix that:

```{r}
par(cex.axis=0.8)
boxplot(hwy ~ class, data = mpg)
```

## 2.2.4 scatterplots







