---
title: "Intro to Econometrics with R Ch. 3"
output: html_notebook
---


The cars dataset is a bunch of cars, their speed, and how long it takes them to stop.

```{r}
View(cars)
```


How big is it?

```{r}
dim(cars)
```

fifty rows, two columns.

## 3.1.2 how are speed and dist related?

WE can start by plotting a graph:

```{r}
plot(dist ~ speed, data = cars,
     xlab = "Speed (in Miles Per Hour)",
     ylab = "Stopping Distance (in Feet)",
     main = "Stopping Distance vs Speed",
     pch  = 20,
     cex  = 2,
     col  = "red")
```

# 3.1 and 3.2 are mostly regression theory

# 3.3 predictions and residuals

```{r}
all.equal(mean(ss$error), 0)
```

Fuck, they don't give you data for several chapters.

Oh well.

# 3.7 a log wage equation

```{r}
data("wage1", package = "wooldridge")   # load data
```

```{r}
# a function that returns a plot
plotfun <- function(wage1,log=FALSE,rug = TRUE){
    y = wage1$wage
    if (log){
        y = log(wage1$wage)
    }
    plot(y = y,
       x = wage1$educ, 
       col = "red", pch = 21, bg = "grey",     
       cex=1.25, xaxt="n", frame = FALSE,      # set default x-axis to none
       main = ifelse(log,"log(Wages) vs. Education, 1976","Wages vs. Education, 1976"),
       xlab = "years of education", 
       ylab = ifelse(log,"Log Hourly wages","Hourly wages"))
    axis(side = 1, at = c(0,6,12,18))         # add custom ticks to x axis
    if (rug) rug(wage1$wage, side=2, col="red")        # add `rug` to y axis
}

par(mfcol = c(2,1))  # set up a plot with 2 panels
# plot 1: standard scatter plot
plotfun(wage1)

# plot 2: add a panel with histogram+density
hist(wage1$wage,prob = TRUE, col = "grey", border = "red", 
     main = "Histogram of wages and Density",xlab = "hourly wage")
lines(density(wage1$wage), col = "black", lw = 2)
```

The distribution of linear wage is skewed, obviously, cos money is like that.


Let's do a first regression of wage vs education:

```{r}
hourly_wage <- lm(formula = wage ~ educ, data = wage1)
plotfun(wage1)
abline(hourly_wage, col = 'black', lw = 2) # add regression line
```

```{r}
summary(hourly_wage)
```


Let's instead do a log regression. 

The update function takes an existing lm object, like hourly_wage here, and updates the formula. Here the . on the right hand side means leave unchanged (so the RHS stays unchanged). 

```{r}
log_hourly_wage = update(hourly_wage, log(wage) ~ ., data = wage1)
```

Now this changes things:

```{r}
par(mfrow = c(1,2))

plotfun(wage1,rug = FALSE)
abline(hourly_wage, col = 'black', lw = 2) # add regression line

plotfun(wage1,log = TRUE, rug = FALSE)
abline(log_hourly_wage, col = 'black', lw = 2) # add regression line
```


```{r}
summary(log_hourly_wage)
```

R-squared is a bit better.


### CEO salaries

Here's another dataset, for CEO salaries:

```{r}
data("ceosal1", package = "wooldridge")  
#
par(mfrow = c(1,2))
#
plot(salary ~ sales, data = ceosal1, main = "Sales vs Salaries",xaxt = "n",frame = FALSE)
axis(1, at = c(0,40000, 80000))
rug(ceosal1$salary,side = 2)
rug(ceosal1$sales,side = 1)
#
plot(log(salary) ~ log(sales), data = ceosal1, main = "Log(Sales) vs Log(Salaries)")
```

```{r}
summary(lm(log(salary) ~ log(sales), data=ceosal1))
```








