---
title: "Intro to Econometrics with R Ch. 2"
output: html_notebook
---

#2.1 Summary statistics

```{r}
library(ggplot2)
```

Let's use that mpg dataset

```{r}
mpg
```


```{r}
mpg$cty
```

We can get mean and median:

```{r}
mean(mpg$cty)
median(mpg$cty)
```

We can get variance, SD, interquartile range, minimum, maximum, and range:

```{r}
var(mpg$cty)
sd(mpg$cty)
IQR(mpg$cty)
min(mpg$cty)
max(mpg$cty)
range(mpg$cty)
```


#2.2 Plotting

You can plot very simple things using R's built-in plot facilities:

```{r}
hist(mpg$cty)
```


And we can pretty up a plot:

```{r}
hist(mpg$cty,
     xlab   = "Miles Per Gallon (City)",
     main   = "Histogram of MPG (City)", # main title
     breaks = 12,   # how many breaks?
     col    = "red",
     border = "blue")
```


For a categorical variable, you can use barplot:

```{r}
barplot(table(mpg$drv),
        xlab   = "Drivetrain (f = FWD, r = RWD, 4 = 4WD)",
        ylab   = "Frequency",
        main   = "Drivetrains",
        col    = "dodgerblue",
        border = "darkorange")
```


You can find out how many unique values are in each column:

```{r}
unique(mpg$drv)
```


You can use a boxplot to visualize a single numerical variable.

The box is the interquartile range, the whiskers are 1.5x IQR, the midline is the median, and the circles are outliers from the IQR.

```{r}
boxplot(mpg$hwy)
```


And you can use this to compare variables:

```{r}
boxplot(hwy ~ class, data = mpg)
```

Note it didn't print 2 classes:

```{r}
unique(mpg$class)
```


The text was too big. 

We can fix that:

```{r}
par(cex.axis=0.8)
boxplot(hwy ~ class, data = mpg)
```

## 2.2.4 scatterplots

To visualize two variables:

```{r}
plot(hwy ~ displ, data = mpg)
```


```{r}
plot(hwy ~ displ, data = mpg,
     xlab = "Engine Displacement (in Liters)",
     ylab = "Miles Per Gallon (Highway)",
     main = "MPG (Highway) vs Engine Displacement",
     pch  = 20,
     cex  = 2,
     col  = "dodgerblue")
```

## 2.2.5 ggplot

You can do all that stuff using ggplot as well:

```{r}
ggplot(data = mpg, mapping = aes(x=displ,y=hwy)) +   
  geom_point(color="blue",size=2) +     # how to show x and y?
  scale_y_continuous(name="Miles Per Gallon (Highway)") +  
  scale_x_continuous(name="Engine Displacement (in Liters)") + 
  theme_bw() +    # change the background
  ggtitle("MPG (Highway) vs Engine Displacement")   # add a title
```

#2.3 Summarizing two variables

You can get covariance easy:

```{r}
cov(mpg$hwy,mpg$displ)
```

And correlation, which is scaleless:

```{r}
cor(mpg$hwy,mpg$displ)

```


# 2.4 The tidyverse

You can do commands with tidy stuff instead:

```{r}
library(dplyr)
```

Here are the most fuel efficient vehicles, filtered:

```{r}
mpg %>% 
  filter(hwy > 35) %>% 
  select(manufacturer, model, year)
```


# 2.4.3 Tidy Example: Importing Non-Tidy Excel Data


```{r}
library(readxl)
```


```{r}
tot_pop_raw = read_excel(
                path = system.file(package="ScPoEconometrics",
                                    "datasets","demo_gind.xls"), 
                sheet="Data", # which sheet
                range="A9:K68")  # which excel cell range to read
names(tot_pop_raw)[1] <- "Country"   # lets rename the first column
tot_pop_raw
```

This shows a tibble, which we encountered just above. The column names are Country,2008,2009,..., and the rows are numbered 1,2,3,.... Notice, in particular, that all columns seem to be of type <chr>, i.e. characters - a string, not a number! We’ll have to fix that, as this is clearly numeric data.

Each year is a column name (like 2008) instead of all years being collected in one column year. We really would like to have several rows for each Country, one row per year. We want to gather() all years into a new column to tidy this up - and here is how:


```{r}
library(tidyr)   # for the gather function
tot_pop = gather(tot_pop_raw, paste(2008:2017),key="year", value = "counts")
tot_pop
```

convert chr to a number for population:

```{r}
tot_pop$counts = as.integer(tot_pop$counts)
```

###dplyr

With dplyr you can do the following operations on data.frames and tibbles:

Choose observations based on a certain value (i.e. subset): filter()
Reorder rows: arrange()
Select variables by name: select()
Create new variables out of existing ones: mutate()
Summarise variables: summarise()
All of those verbs can be used with group_by(), where we apply the respective operation on a group of the dataframe/tibble. For example, on our tot_pop tibble we will now

filter
mutate
and plot the resulting values
Let’s get a plot of the populations of France, the UK and Italy over time, in terms of millions of people. We will make use of the piping syntax of dplyr which we introduced just above.


```{r}
library(dplyr)  # for %>%, filter, mutate, ...
# 1. take the data.frame `tot_pop`
tot_pop %>%
  # 2. pipe it into the filter function
  # filter on Country being one of "France","United Kingdom" or "Italy"
  filter(Country %in% c("France","United Kingdom","Italy")) %>%
  # 3. pipe the result into the mutate function
  # create a new column called millions
  mutate(millions = counts / 1e6) %>%
  # 4. pipe the result into ggplot to make a plot
  ggplot(mapping = aes(x=year,y=millions,color=Country,group=Country)) + geom_line(size=1)
```

How to find all NAs in a tibble:

```{r}
missings = tot_pop %>%
  filter(is.na(counts)) %>%    # is.na(x) returns TRUE if x is NA
  group_by(Country) %>%
  summarise(n_missing = n(),years = paste(year,collapse = ", "))
knitr:::kable(missings)     # knitr:::kable makes a nice table
```









